---
title: "p8105_hw3_wc2750"
author: "Weixi Chen"
date: "10/9/2020"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(readxl)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d

```

## Problem 1
Load instacart dataset
```{r}
library(p8105.datasets)
data("instacart")
```

This dataset contains `r nrow(instacart)` rows and `r ncol(instacart)` columns.

Observations are the level of items in orders by user. There are user / order variables -- user ID, order ID, order day, and order hour. There are also item variables -- name, aisle, department, and some numeric codes.

How many aisles, and which are most items from?
```{r}
instacart %>%
  count(aisle) %>%
  arrange(desc(n))
```

Let's make a plot
```{r}
instacart %>%
  count(aisle) %>%
  filter(n > 10000) %>%
  mutate(
    aisle = factor(aisle),
    aisle = fct_reorder(aisle, n)
  ) %>%
  ggplot(aes(x= aisle, y = n)) +
  geom_point() +
  labs(y = "number of items", title = "Number of items ordered in each aisle") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1), plot.title = element_text(hjust = 0.5))
```

Let's make a table
```{r}
instacart %>%
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>%
  group_by(aisle) %>%
  count(product_name) %>%
  mutate(rank = min_rank(desc(n))) %>%
  filter(rank <4) %>%
  arrange(aisle, rank) %>%
  knitr::kable()
```

Apple vs ice cream
```{r message = FALSE}
instacart %>%
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>%
  group_by(product_name, order_dow) %>%
  summarize(mean_hour = mean(order_hour_of_day)) %>%
  pivot_wider(
    names_from = order_dow,
    values_from = mean_hour
  )
```

## Problem 2
### Part 1
Import and tidy accel_data
```{r message = FALSE}
accel_df = read_csv("./data/accel_data.csv") %>%
  janitor::clean_names() %>%
  pivot_longer(
    cols = starts_with("activity_"),
    names_to = "minute",
    names_prefix = "activity_",
    values_to = "activity_count"
  ) %>%
  # include a weekday vs weekend variable
  mutate (
    weekday_vs_weekend = ifelse(day %in% c("Saturday", "Sunday"), "weekend", "weekday")
  ) %>%
  # encode data with reasonable variable classes
  mutate(
    week = as.integer(week),
    day_id = as.integer(day_id),
    day = ordered(as.factor(day), levels=c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday")),
    minute = as.integer(minute),
    weekday_vs_weekend = as.factor(weekday_vs_weekend)
  ) %>%
  # arrange in reasonable order and organize variables
  arrange(week,day,minute) %>%
  relocate(week, day_id, day, weekday_vs_weekend)

# preview accel_df
accel_df
```

The accel_df has `r nrow(accel_df)` rows and `r ncol(accel_df)` columns. The observation data is collected from `r max(accel_df$week)` weeks and `r max(accel_df$day_id)` days. The day variable indicates the day of the week. The added weekday_vs_weekend variable specifies whether that day is weekday or weekend. The minute variable ranges from `r min(accel_df$minute)` to `r max(accel_df$minute)`. The activity_counts variable indicates the activity counts for each minute.

### Part 2
Total acitivity over the day
```{r message = FALSE}
# table for aggregate daily activity counts
accel_df %>%
  group_by(day_id) %>%
  summarize(daily_activity = sum(activity_count))
```

```{r message = FALSE}
# table more easier to read
accel_df %>%
  group_by(week,day) %>%
  summarize(daily_activity = sum(activity_count)) %>%
  pivot_wider(
    names_from = day,
    values_from = daily_activity
  )
```

Visualize the trends
```{r message = FALSE}
accel_df %>%
  group_by(week,day) %>%
  summarize(daily_activity = sum(activity_count)) %>%
  ggplot(aes(x = day, y = daily_activity, group = week, color = week)) +
  geom_point() +
  geom_line() +
  labs(x = "day of the week", y = "daily acitivity", title = "Daily activity of each day of a week among the 5-week observation", caption = "data collected on a 63 year-old male with BMI 25") +
  theme(plot.title = element_text(size = 10, hjust = 0.5))
```

For week 1,2, and 5, the activity count of the observed subject is gradually increasing through the weekdays. For week 3, the subject has the maximum activity on Monday and then keeps a stable amount of activity in the following days. For week 4, the activity count keeps stable from Monday to Wednesday but decreases after Wednesday. For week 1, 2, and 3, the activity count during weekends is close to the count on the same week's Friday. For week 4 and 5, the subject has nearly zero amount of activity on Saturday and relatively lower amount on Sunday.

### Part 3
Make a single-panel plot that shows the 24-hour activity time courses for each day and use color to indicate day of the week
```{r message = FALSE}
accel_df %>%
  group_by(day, minute) %>%
  summarize(weekday_activity = sum(activity_count)) %>%
  ggplot(aes(x = minute, y = weekday_activity, color = day)) +
  geom_point(alpha = .3, size = .5) +
  geom_smooth(size = .5) +
  labs(x= "hours", y= "activity count", title = "24-hour activity time courses for each day", caption = "data collected on a 63 year-old male with BMI 25", color = "day of the week") +
  scale_x_continuous(
    breaks = c(seq(0, 1440, by = 60)),
    labels = c(seq(0, 24, by = 1))) +
  theme(plot.title = element_text(hjust = 0.5))
```

The observed subject might sleep around 11:30PM and get up around 5AM. He has relatively higher amount of activity from 10AM to 12PM on Sunday and from 8PM to 10PM on Friday. For the rest of the time, his activity amount keeps stable.
